### 2022-10-16 main.tex
added abstract/conclusion; fixed results a bit

%%%%%%%%%%%%%%%%%%%%%%% file template.tex %%%%%%%%%%%%%%%%%%%%%%%%%
%
% This is a general template file for the LaTeX package SVJour3
% for Springer journals.          Springer Heidelberg 2010/09/16
%
% Copy it to a new file with a new name and use it as the basis
% for your article. Delete % signs as needed.
%
% This template includes a few options for different layouts and
% content for various journals. Please consult a previous issue of
% your journal as needed.
%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%
% First comes an example EPS file -- just ignore it and
% proceed on the \documentclass line
% your LaTeX will extract the file if required
%
\RequirePackage{fix-cm}
%
%\documentclass{svjour3}                     % onecolumn (standard format)
%\documentclass[smallcondensed]{svjour3}     % onecolumn (ditto)
\documentclass[smallextended]{svjour3}       % onecolumn (second format)
%\documentclass[twocolumn]{svjour3}          % twocolumn
%
\smartqed  % flush right qed marks, e.g. at end of proof
%
\usepackage{graphicx}
%
% \usepackage{mathptmx}      % use Times fonts if available on your TeX system
%
% insert here the call for the packages your document requires
%\usepackage{latexsym}
% etc.
%
% please place your own definitions here and don't use \def but
% \newcommand{}{}
%
% Insert the name of "your journal" with
\journalname{Empirical Software Engineering}
%

\usepackage{listings, listings-rust}
\lstset{escapeinside={<@}{@>}}
\usepackage{hyperref}
\usepackage{tabularx}
\usepackage{array}
\usepackage{graphicx}
\usepackage{algorithm}
\usepackage{algpseudocode}
\usepackage[utf8]{inputenc}
\usepackage{tikz}
\usetikzlibrary{graphs} 
\usetikzlibrary{matrix}
\usepackage{multirow}
% \usepackage[backend=bibtex, style=authoryear]{biblatex}
% \addbibresource{bibtex.bib}
\usepackage{subfiles} % Best loaded last in the preamble
\input{defs}

\tikzset{ 
    table/.style={
        matrix of nodes,
        row sep=-\pgflinewidth,
        column sep=-\pgflinewidth,
        nodes={
            rectangle,
            draw=black,
            align=center
        },
        minimum height=1.5em,
        text depth=0.5ex,
        text height=2ex,
        nodes in empty cells,
        column 1/.style={
            nodes={text width=5em,font=\bfseries, fill=gray!20}
        },
        column 2/.style={
            nodes={text width=4em,font=\bfseries}
        },
        column 3/.style={
            nodes={text width=5em,font=\bfseries}
        },
        column 4/.style={
            nodes={text width=5em,font=\bfseries}
        },  
        column 5/.style={
            nodes={text width=8em,font=\bfseries}
        },  
        column 6/.style={
            nodes={text width=5em,font=\bfseries}
        },  
        column 7/.style={
            nodes={text width=5em,font=\bfseries}
        },
        column 8/.style={
            nodes={text width=9em,font=\bfseries}
        },
        column 9/.style={
            nodes={text width=3em,font=\bfseries}
        }
    }
}

\begin{document}

\title{A Study on Common Bug Fix Patterns in Rust%\thanks{Grants or other notes
%about the article that should go on the front page should be
%placed here. General acknowledgments should be placed at the end of the article.}
}
\subtitle{}

%\titlerunning{Short form of title}        % if too long for running head

\author{Mohammad Robati Shirzad         \and
        Patrick~Lam %etc.
}

%\authorrunning{Short form of author list} % if too long for running head

\institute{F. Author \at
              first address \\
              Tel.: +123-45-678910\\
              Fax: +123-45-678910\\
              \email{fauthor@example.com}           %  \\
%             \emph{Present address:} of F. Author  %  if needed
           \and
           S. Author \at
              second address
}

\date{Received: date / Accepted: date}
% The correct dates will be entered by the editor

\maketitle

\begin{abstract}
    Rust is a relatively new programming language that allows programmers to write programs that have low-level control over resources, as well as high-level safety guarantees. This dual nature is made possible through the ownership framework which has been incorporated in the language. The main component governing over this framework is called Borrow-Checker. Rust has been ranked first for the sixth consecutive year in the Stack Overflow's annual Developer Survey as the most loved programming language. The number of Rust developers is growing as the need for faster and safer software increases. However, to the best of our knowledge, no research has been developed to automatically find the most pervasive bug fix patterns within Rust programs. In this project, using a novel program modelling method, we mined the top 18 most starred Rust projects in Github to discover the most common bug fix patterns lied within their code changes. We stored 87,726 code changes drawn from 57,214 commits across these 18 projects. After clustering the code changes, and conducting a manual analysis, we discovered 19 groups of cross-project bug fix patterns. We introduce these patterns in two groups: (1) general patterns and (2) borrow-checker related patterns.
\keywords{First keyword \and Second keyword \and More}
% \PACS{PACS code1 \and PACS code2 \and more}
\subclass{MSC code1 \and MSC code2 \and more}
\end{abstract}

\section{Introduction}
% rubber: depend sections/introduction.tex
\subfile{sections/introduction}

\section{Methodology}
% rubber: depend sections/methodology.tex
\subfile{sections/methodology}

\section{Results}
% rubber: depend sections/results.tex
\subfile{sections/results}

\section{Discussion}
% rubber: depend sections/discussion.tex
\subfile{sections/discussion}

\section{Related Work}
% rubber: depend sections/related.tex
\subfile{sections/related}

\section{Conclusion}
% rubber: depend sections/conclusion.tex
\subfile{sections/conclusion}

%\begin{acknowledgements}
%If you'd like to thank anyone, place your comments here
%and remove the percent signs.
%\end{acknowledgements}

% Authors must disclose all relationships or interests that 
% could have direct or potential influence or impart bias on 
% the work: 
%
% \section*{Conflict of interest}
%
% The authors declare that they have no conflict of interest.

% BibTeX users please use one of
\bibliographystyle{plain}      % basic style, author-year citations
% \bibliographystyle{spmpsci}      % mathematics and physical sciences
% \bibliographystyle{spphys}       % APS-like style for physics
\bibliography{refs}       % name your BibTeX data base
% \printbibliography[title={References}]  % name your BibTeX data base

% Non-BibTeX users please use
% \begin{thebibliography}{}
% %
% % and use \bibitem to create references. Consult the Instructions
% % for authors for reference list style.
% %
% \bibitem{RefJ}
% % Format for Journal Reference
% Author, Article title, Journal, Volume, page numbers (year)
% % Format for books
% \bibitem{RefB}
% Author, Book title, page numbers. Publisher, place (year)
% % etc
% \end{thebibliography}

\section{Appendix}
\subfile{sections/appendix}

\end{document}
% end of file template.tex