### 2022-10-16 sections/conclusion.tex

In this project, we mined the top 18 most starred Rust projects in Github to discover common bug fix patterns lied within their code changes. We proposed a novel program modelling approach to encode most important aspects of a code change in a fixed sized datapoint. We stored 87,726 datapoints drawn from 57,214 commits across these 18 projects. After using the DBSCAN clustering algorithm, and a subsequent manual analysis, we obtained 10 clusters of general fix patterns and 9 clusters of borrow-checker related fix patterns.