### 2022-11-02 conclusion.tex
added Ruxanne as the name

In this project, we developed Ruxanne, a tool for mining bug fix patterns in Rust. Using this tool, we mined the top 18 most starred Rust projects in Github to discover common bug fix patterns lied within their code changes. At the heart of Ruxanne, we used a novel code embedding approach to embed most important aspects of a code change in a fixed sized datapoint. We stored 87,726 datapoints drawn from 57,214 commits across these 18 projects. After using the DBSCAN clustering algorithm, and a subsequent manual analysis, we obtained 12 clusters of general fix patterns and 8 clusters of borrow-checker related fix patterns.