### 2023-10-31 response.tex
EMSE revision#2 - reviewer comments added to response.tex, todos added, two of the raised minor issues addressed and fixed

\documentclass{article}

\usepackage{cite}
\usepackage{graphicx}
\usepackage{hyperref}
\usepackage{todonotes}

\title{Response to Reviewer Comments}
\date{2023 November}
\author{Mohammad Robati Shirzad and Patrick Lam}
\begin{document}

\maketitle

Please find below our response to the reviews for our EMSE submission ``A Study of Common Bug Fix Patterns in Rust''.

\section{Reviewer 1}

\textit{I appreciate the authors' efforts in improving the paper, and most of my concerns have been addressed.}

\subsection{Reorganizing related work}

\textit{\textbf{R1:} I have observed that the related work now includes several subsections focusing on APR and patch correctness prediction. It would be beneficial to reorganize and consolidate these subsections into a single coherent section. Furthermore, the section on related work should encompass studies related to bug classification in programming languages with a longer history.}

\vspace*{1em} \noindent \textbf{Response:} \todo{Probably no need for argument, let's just do what he asks.}

\subsection{More discussion on the impact of the evolving Rust specification}

\textit{\textbf{R1:} Additionally, I would appreciate it if the authors could discuss the potential impact of the evolving Rust specification. Given that Rust is a novel programming language, it is subject to rapid grammatical changes, which in turn could influence bug characteristics and symptoms.}

\vspace*{1em} \noindent \textbf{Response:} \todo{Probably no need for argument, let's just do what he asks.}

\section{Reviewer 2}

\textit{Thanks a lot for the information and clarification in the rebuttal and the new paper version. I would suggest the authors make the following changes in the next paper version.}

\subsection{More discussion on the size and type of the patches}

\textit{\textbf{R2:} Inspired by the last paragraph of Section 4.1, can authors discuss which of the following is more important for Rust in the next paper version: fixing bugs and fixing compilation errors?}

\vspace*{1em} \noindent \textbf{Response:} \todo{Need some argument here and some changes in the text.}

\vspace*{1em} \noindent \textit{\textbf{R2:} Please discuss why identified patches are all small in the next paper version. Basically, please add the last paragraph of Section 2.2.4 of the rebuttal to the next paper version.}

\vspace*{1em} \noindent \textbf{Response:} \todo{Probably just do what he asks.}

\subsection{Paraphrasing and providing clearer explanation}

\textit{\textbf{R2:} In Section 2.3.2, the authors use sentences from an existing paper directly (although quotes are added). I don't think this is a good way to write an academic paper. Please paraphrase these sentences.}

\vspace*{1em} \noindent \textbf{Response:} \todo{No argument, just paraphrase.}

\vspace*{1em} \noindent \textit{\textbf{R2: } I still do not understand Figure 1. I suggest the authors label ASTDiff nodes with code to ease the explanation.}

\vspace*{1em} \noindent \textbf{Response:} \todo{First, we need some changes in paper text to make it clearer. Then, need some explanation here how we've changed it.}

\subsection{Minor issues}

\textit{\textbf{R2:} Line 21 on page 8: work around -> workaround}

\vspace*{1em} \noindent \textbf{Response:} Fixed.

\vspace*{1em} \noindent \textit{\textbf{R2:} Line 49 on page 8: which path is colored in purple in figure 1?}

\vspace*{1em} \noindent \textbf{Response:} It is the rightmost path. We changed the color to blue so it is clearer.

\small
\bibliographystyle{plain}
\bibliography{refs}

\end{document}