### 2023-11-09 response.tex
reviewer 2 finished for now

\documentclass{article}

\usepackage{cite}
\usepackage{graphicx}
\usepackage{hyperref}
\usepackage{todonotes}

\title{Response to Reviewer Comments}
\date{2023 November}
\author{Mohammad Robati Shirzad and Patrick Lam}
\begin{document}

\maketitle

Please find below our response to the reviews for our EMSE submission ``A Study of Common Bug Fix Patterns in Rust''.

\section{Reviewer 1}

\textit{I appreciate the authors' efforts in improving the paper, and most of my concerns have been addressed.}

\subsection{Reorganizing related work}

\textit{\textbf{R1:} I have observed that the related work now includes several subsections focusing on APR and patch correctness prediction. It would be beneficial to reorganize and consolidate these subsections into a single coherent section. Furthermore, the section on related work should encompass studies related to bug classification in programming languages with a longer history.}

\vspace*{1em} \noindent \textbf{Response:} \todo{Probably no need for argument, let's just do what he asks.}

\subsection{More discussion on the impact of the evolving Rust specification}

\textit{\textbf{R1:} Additionally, I would appreciate it if the authors could discuss the potential impact of the evolving Rust specification. Given that Rust is a novel programming language, it is subject to rapid grammatical changes, which in turn could influence bug characteristics and symptoms.}

\vspace*{1em} \noindent \textbf{Response:} \todo{Probably no need for argument, let's just do what he asks.}

\section{Reviewer 2}

\textit{Thanks a lot for the information and clarification in the rebuttal and the new paper version. I would suggest the authors make the following changes in the next paper version.}

\subsection{More discussion on the size and type of the patches}

\textit{\textbf{R2:} Inspired by the last paragraph of Section 4.1, can authors discuss which of the following is more important for Rust in the next paper version: fixing bugs and fixing compilation errors?}

\vspace*{1em} \noindent \textbf{Response:} Added more discussion regarding this to the last paragraph of Section 4.1. We discuss that we consider bugs as potentially more critical, because an unresolved bug can remain dormant within the system and unintentionally get deployed, whereas a program with compilation errors cannot be shipped by developers.

\vspace*{1em} \noindent \textit{\textbf{R2:} Please discuss why identified patches are all small in the next paper version. Basically, please add the last paragraph of Section 2.2.4 of the rebuttal to the next paper version.}

\vspace*{1em} \noindent \textbf{Response:} Added the mentioned paragraph to the Discussion seciton of the manuscript.

\subsection{Paraphrasing and providing clearer explanation}

\textit{\textbf{R2:} In Section 2.3.2, the authors use sentences from an existing paper directly (although quotes are added). I don't think this is a good way to write an academic paper. Please paraphrase these sentences.}

\vspace*{1em} \noindent \textbf{Response:} Paraphrased the three definitions that were directly quoted.

\vspace*{1em} \noindent \textit{\textbf{R2: } I still do not understand Figure 1. I suggest the authors label ASTDiff nodes with code to ease the explanation.}

\vspace*{1em} \noindent \textbf{Response:} Labeled ASTDiff nodes with code. \todo{Should we add more explanation?}

\subsection{Minor issues}

\textit{\textbf{R2:} Line 21 on page 8: work around -> workaround}

\vspace*{1em} \noindent \textbf{Response:} Fixed.

\vspace*{1em} \noindent \textit{\textbf{R2:} Line 49 on page 8: which path is colored in purple in figure 1?}

\vspace*{1em} \noindent \textbf{Response:} It is the rightmost path. We changed the color to blue so it is clearer.

\small
\bibliographystyle{plain}
\bibliography{refs}

\end{document}